% SPDX-License-Identifier: AGPL-3.0-or-later
% Compile: pdflatex 02-arxiv-note.tex  (run twice for references/labels)
% Self-contained: inline thebibliography, no .bib required.
\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage[hidelinks]{hyperref}
\usepackage{titlesec}
\titleformat*{\section}{\large\bfseries}
\titleformat*{\subsection}{\normalsize\bfseries}

\newcommand{\safe}{\texttt{safe}}
\newcommand{\warning}{\texttt{warning}}
\newcommand{\blocked}{\texttt{blocked}}

\title{\textbf{A Frozen-Benchmark Methodology for Measured\\ Prompt-Injection Boundary Hardening}}
\author{Raouf\\ \texttt{raoof.r12@gmail.com}}
\date{2026-06-16}

\begin{document}
\maketitle

\begin{abstract}
\noindent
We present an application-layer boundary that classifies user input before a language
model is invoked and emits a tamper-evident, metadata-only receipt for every decision,
together with a methodology that makes boundary hardening \emph{measurable} rather than
asserted. The difficulty is methodological as much as technical: prompt-injection filters
are easy to improve on their own test cases and hard to evaluate honestly, because the
corpus, the success criterion, and the failure baseline are usually written by the same
hand that reports the win. We address this by freezing an adversarial corpus and its
cryptographic detector digests, publishing the seed detector's honest failure baseline
($2/30$) before hardening, and adopting a success criterion of \emph{strict improvement
under fixed guardrails} rather than a target number. A deterministic
canonicalize-then-classify pipeline with a framing-aware context guard and a three-tier
verdict (\safe/\warning/\blocked) then improves adversarial detection from $2/30$ to
$\mathbf{18/30}$ on the frozen corpus while preserving clean-benign pass at $10/10$ and
\emph{reducing} hard-negative blocked false positives from $2/5$ to $0/5$; on a held-out
set authored after the detector froze, detection generalizes to $7/9$. The most useful
number is the one we did not optimize: the context guard adds zero detection and exists
solely to cut false positives.
\end{abstract}

\noindent\textit{Compact technical note (Stages 3A--3C of Project Simurgh LLM Shield).
This is a methodology-and-evaluation note, not a claim of a new state-of-the-art
classifier.}

\section{Introduction}
When Anthropic suspended access to two of its most capable models in June 2026, it stated
that it suspects ``perfect jailbreak resistance is not currently possible for any model
provider,'' and described its strategy as defence-in-depth combined with monitoring
\cite{anthropic2026}. The security-engineering reaction emphasized guardrails over kill
switches, insistence on evidence and a remediation path, and treating AI systems as
attack surface under the OWASP LLM Top 10 \cite{snyk2026,owasp2025}. Both point to the
same gap: the application layer needs to \emph{prove what its safety boundary did} on each
input, not claim the boundary is perfect.

We make three contributions, in increasing order of importance:
\begin{itemize}
  \item \textbf{A boundary with auditable evidence.} An input-only, pre-provider
  classifier that emits HMAC-chained, metadata-only receipts for \safe/\warning/\blocked{}
  decisions.
  \item \textbf{A deterministic hardening pipeline.} Canonicalize-then-classify (homoglyph
  fold, leet/symbol de-stuffing, base64 decode-for-inspection) plus a framing-aware
  context guard that separates \emph{recall} mechanisms from \emph{precision} mechanisms.
  \item \textbf{A freeze-then-harden methodology.} A frozen corpus with digest-pinned
  detectors, a published failure baseline, a no-target success criterion, an ablation, and
  a held-out generalization probe --- designed so a measured improvement cannot silently
  become a goalpost move.
\end{itemize}

\section{Threat model and boundary}
We protect the boundary \emph{before} the provider is invoked: preventing an obfuscated
policy-override or system-prompt-exfiltration instruction from reaching the model
unflagged, and producing tamper-evident proof of the decision. The in-scope adversary is
single-turn and fully controls the input string, obfuscating via Unicode homoglyphs,
symbol/space stuffing, base64, role-play framing, structured (JSON/Markdown) hidden
channels, or translate-then-follow wrappers. Out of scope (and deferred to later stages):
untrusted retrieved context, tool/function-call abuse, live-model-specific exploits, and
multi-turn memory poisoning. The audit secret is assumed not attacker-known; raw payloads
exist only in fixtures, and all generated evidence is metadata-only.

\section{Method}
\paragraph{Canonicalization.} From the normalized input we derive two views. The
\emph{canonical} view folds a small, audited table of Unicode confusables to ASCII and
applies leet/symbol substitutions in word context ($\mathtt{!|}\rightarrow\mathtt{i}$,
$\mathtt{1}\rightarrow\mathtt{i}$, $\mathtt{0}\rightarrow\mathtt{o}$, \dots), leaving
digits without letter neighbours untouched. Base64-looking blobs are decoded \emph{before}
folding (base64 is case-sensitive and leet-folding would corrupt blob digits) and appended
for inspection only --- never executed. The \emph{compact} view strips all
non-alphanumerics, so separator- and symbol-stuffed attacks rejoin to their real
spellings. Each transformation that fires is recorded as an enum signal (e.g.\
\texttt{homoglyph\_fold}, \texttt{symbol\_destuff}, \texttt{base64\_decoded}), never as raw
text.

\paragraph{Classification.} A negation-aware denylist is matched over both views, augmented
with deterministic heuristics for role-play framing, structured hidden-instruction
channels, and translate-then-follow wrappers. This design is \emph{consistent with} OWASP
LLM01:2025's input/output filtering and string-checking guidance and directly targets its
documented encoded/multilingual scenario \cite{owasp2025}; Unicode-injection
canonicalization is recognized industry prior art \cite{patent}. We claim no novelty for
canonicalization itself.

\paragraph{Context guard and verdicts.} A framing-aware \emph{deterministic} guard
(heuristics, not semantic understanding) de-escalates a block-worthy match to \warning{}
only when the \emph{matched} attack phrase is itself quoted or the utterance carries an
educational lead-in; a stray quoted word elsewhere cannot soften a bare attack. Verdicts:
\blocked{} (denylist match outside educational framing; provider skipped), \warning{}
(weaker/contextual signals, or a quoted/educational match; provider still called, warning
receipt emitted), \safe{} (no signal). A three-class output reflects that not every
instruction-looking input is malicious, consistent with recent detection work
\cite{promptsleuth}.

\section{Evaluation}
\paragraph{Corpus (frozen).} 30 adversarial cases across 10 obfuscation styles (3 each)
and 15 benign cases (5 normal, 5 AI-safety questions, 5 hard negatives that quote attack
phrases in educational framing). Each fixture stores \texttt{ground\_truth} (what the case
\emph{is}) and a separately re-snapshotted \texttt{baseline\_verdict} (what the detector
\emph{does}); detector source files are pinned by SHA-256 digests. The corpus payloads are
never edited during hardening --- only the recorded baseline is re-snapshotted through a
reviewed update path.

\paragraph{Success criterion.} Strictly improve adversarial detection above the frozen
$2/30$ baseline while preserving clean-benign pass at $10/10$ and holding hard-negative
blocked false positives $\leq 2/5$. No fixed target is set, to avoid detector-gaming.
Detection counts \blocked{} \emph{or} \warning{} (a warning on an attack is a catch); a
benign false positive counts \blocked{} only (a warning on a benign is acceptable
friction).

\begin{table}[h]
\centering
\caption{Headline result on the frozen corpus.}
\begin{tabular}{lcc}
\toprule
Metric & Seed (3B) & Hardened (3C) \\
\midrule
adversarial detection (\blocked{} $\vee$ \warning) & $2/30$ & $\mathbf{18/30}$ \\
\quad --- blocked / warning & $2\,/\,0$ & $13\,/\,5$ \\
clean-benign pass & $10/10$ & $\mathbf{10/10}$ \\
hard-negative blocked FP & $2/5$ & $\mathbf{0/5}$ \\
\bottomrule
\end{tabular}
\end{table}

We note honestly that the seed's two detections were incidental: their second clause
matched the denylist while the obfuscated first clause was unhandled. Part of the 3C gain
is also a larger denylist (spaced matching alone reaches $5/30$); the pipeline's
contribution is the increment above that, isolated by the ablation.

\begin{table}[h]
\centering
\caption{Ablation: per-stage contribution on the frozen corpus.}
\begin{tabular}{lccc}
\toprule
Configuration & detection & clean & hard-neg FP \\
\midrule
spaced phrase-match only & $5/30$ & $10/10$ & $2/5$ \\
\quad + canonicalization & $14/30$ & $10/10$ & $2/5$ \\
\quad + heuristics & $18/30$ & $10/10$ & $2/5$ \\
\quad + context guard (full) & $18/30$ & $10/10$ & $\mathbf{0/5}$ \\
\bottomrule
\end{tabular}
\end{table}

Canonicalization is responsible for most of the recall gain; the context guard adds no
detection and improves only precision. The two mechanisms are decoupled and individually
legible.

\paragraph{Held-out generalization.} On 12 new surface forms authored after the detector
froze and never used to tune it, adversarial detection is $7/9$, clean $2/2$, hard-neg FP
$0/1$. The two misses are \texttt{academic-framing} and \texttt{multi-step-softening} ---
the same semantic styles that miss on the frozen corpus --- indicating a genuine
capability ceiling rather than corpus overfitting.

\section{Related work}
Our position is that of an \emph{evidence/audit layer}, not a competing classifier.
PromptSleuth argues detection should key on semantic intent because surface attack forms
vary widely, which supports our explicit non-claim that phrase/canonical matching is
incomplete \cite{promptsleuth}. WAInjectBench shows detectors struggle in web-agent
settings, supporting corpus-anchored, measured evaluation over blanket efficacy claims
\cite{wainjectbench}. Three-class framings in recent detection work support our
\safe/\warning/\blocked{} tier \cite{alignsentinel}. OWASP LLM01:2025 documents the
layered defences and encoded/multilingual scenarios we target \cite{owasp2025}, and
Unicode-injection canonicalization is established prior art \cite{patent}.

\section{Limitations}
\textbf{Deterministic ceiling.} Styles whose malice lives in intent rather than surface
form --- \texttt{multi-step-softening} and most \texttt{academic-framing} --- are largely
unreachable by canonicalization plus phrase matching, on both the frozen and held-out
sets. \textbf{Statistical power.} $n=30$ adversarial / $15$ benign ($+12$ held-out) is
illustrative, not statistically powered; results are fractions, not population-level
claims. \textbf{Canonicalization is attackable.} The fold/de-stuff tables are finite; an
adversary aware of them can craft around them, which is why each future bypass becomes a
frozen regression fixture. \textbf{No semantic understanding.} The context guard is
framing-aware heuristics; unusual benign framings can still be misread.

\section{Conclusion}
The contribution is a posture, demonstrated end to end: publish the failure, freeze the
evidence, harden without moving the goalposts, and keep every decision auditable. Against
a frozen corpus this turned $2/30$ into $18/30$ with no loss of benign utility and a
reduction in false positives, with a held-out probe to estimate overfitting. The same
methodology extends to the harder boundaries --- untrusted context, tool use, live models
--- which are the natural next stages.

\paragraph{Non-claims.} We do not claim jailbreak immunity or provider-level model safety;
no live model is used; untrusted context, tools, and multi-turn memory are out of scope.
Detection is deterministic and incomplete by construction. Receipts attest process, not
ground truth.

\paragraph{Reproducibility.}
\texttt{scripts/smoke-llm-shield-bench.sh} (headline; asserts no drift),
\texttt{tests/e2e/llm\_shield\_ablation\_runner.mjs} (ablation),
\texttt{tests/e2e/llm\_shield\_heldout\_runner.mjs} (held-out).

\begin{thebibliography}{9}
\bibitem{anthropic2026} Anthropic. \emph{Statement on the US government directive to
suspend access to Fable 5 and Mythos 5.}
\url{https://www.anthropic.com/news/fable-mythos-access}
\bibitem{snyk2026} Snyk. \emph{When a Government Pulls an AI Model: takeaways for security
teams.} \url{https://snyk.io/blog/fable-mythos-suspension-security-takeaways/}
\bibitem{owasp2025} OWASP Gen AI Security Project. \emph{LLM01:2025 Prompt Injection.}
\url{https://genai.owasp.org/llmrisk/llm01-prompt-injection/}
\bibitem{patent} \emph{Canonicalization of unicode prompt injections} (US patent filing;
cited as industry prior art, not efficacy evidence).
\url{https://patents.justia.com/inventor/jason-martin}
\bibitem{promptsleuth} M. Wang, Y. Zhang, and G. Gu. \emph{PromptSleuth: Detecting Prompt
Injection via Semantic Intent Invariance.} arXiv:2508.20890, 2025.
\url{https://arxiv.org/abs/2508.20890}
\bibitem{wainjectbench} Y. Liu, R. Xu, X. Wang, Y. Jia, and N. Z. Gong.
\emph{WAInjectBench: Benchmarking Prompt Injection Detections for Web Agents.}
arXiv:2510.01354, 2025. \url{https://arxiv.org/abs/2510.01354}
\bibitem{alignsentinel} Y. Jia, R. Wang, X. Wang, C. Xiang, and N. Gong.
\emph{AlignSentinel: Alignment-Aware Detection of Prompt Injection Attacks.}
arXiv:2602.13597, 2026. \url{https://arxiv.org/abs/2602.13597}
\end{thebibliography}

\end{document}
